Forecasting Multi-District Elections

### 2019-11-26 outputs/paper/paper.Rmd
Update paper

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "Forecasting Multi-District Elections"
thanks: "We thank the Social Research Centre, and in particular Ben Phillips, for graciously providing data. Thank you to Monica Alexander, Lauren Kennedy, John McAndrews, Benjamin Stevens, Michael Donnelly, as well as seminar participants at the University of Toronto Political Behaviour Group and the 2019 POP AusPSA Standing Group Annual Workshop for their assistance and helpful suggestions. Our code and datasets are available at: https://github.com/RohanAlexander/ForecastingMultiDistrictElections. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Rohan Alexander
  affiliation: University of Toronto
- name: Patrick Dumont
  affiliation: Australian National University
- name: Patrick Leslie
  affiliation: Australian National University
abstract: "The overall outcome of an election in a multi-district system turns on the number of electoral divisions won by each party. Despite this, traditional political opinion polls tend to focus on broad national-level estimates of support for major parties. In this paper we develop a framework to characterise a useful forecasting model of multi-district elections. We then implement an approach to forecast the number of electoral divisions won by each of the larger political parties in Australia at the 2019 Federal Election. Our approach first uses survey data and multi-level regression with post-stratification (MRP) to estimate electoral-division-level first-preference shares. To account for Australia's use of preferential voting, we then use a model that we trained to estimate electoral-division-level two-party-preferred shares. We find that our approach performs well out-of-sample, and it has additional advantages over traditional polling including improved interpretability, transparency, and better communication of statistical uncertainty. Our framework allows the consistent evaluation of forecasting models of elections in a multi-district system. Finally, our paper also improves our understanding of the circumstances in which MRP is appropriate, as we find that despite the bias of voter advice applications, and despite the relatively small number of respondents per electoral division for usually political opinion polls, both can still provide informative results on an electoral division basis."
keywords: "Multi-district elections, political opinion polls, forecasting, multi-level regression with post-stratification, Australia."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: references.bib
biblio-style: "apalike"
endnote: no

@@

In this paper we show that it is possible to use the individual-level responses already collected by political opinion polling firms, along with publicly available datasets and accessible quantitative methods to produce estimates of the number of electoral divisions won by each party. We first establish a framework for evaluating a forecasting model for multi-district elections. It consists of five aspects. These are that a model should: be interpretable and provide reasonable estimates of the number of electoral divisions that each party will win; account for preferences and/or turnout; operate in a reproducible and transparent way; produce results that account for the fact that it is a probability model; and finally not be worse than traditional approaches in terms of speed or cost. Within this framework, the method that we implement involves first estimating the first-preference share of each party in each division based on survey data using multi-level regression with post-stratification. We then use train a model to estimate division-level two-party-preferred shares.
<!-- Seat-specific anomalies can be accounted for by adding additional layers to the model, or by inputting detailed information about that seat. And the model, which is centered around multi-level regression with post-stratification goes some way to ensuring that a handful of odd results do not need to be manually adjusted or removed. -->

Multi-level regression with post-stratification (MRP) [@GelmanLittle1997] is a popular way to analyse opinion by way of estimating good-quality estimates of subsets of the population. It uses a regression model to consider individual-level survey responses as being determined by various sub-groups and then rebuilds the sample to better match the population. In this way MRP can not only allow a better understanding of opinion, but also allow us to analyse data that may otherwise be unusable and consider outcomes within the context of distributions and uncertainty. MRP has been used in a variety of settings and one of its advantages is that a survey analysed in this way can be biased to a larger extent than normal because the approach should go some way to account for this. For instance, in the US @WangRothschildGoelGelman2015 used data collected from the XBox platform, which tends to be both younger and more male than the US electorate, and were able to draw meaningful insights about the 2012 US Presidential Election.

In this paper we construct a Bayesian MRP model to generate division-level two-party-preferred estimates for the Australian 2019 Federal Election. In our regression model an individual's first-preference is a function of the demographic and geographic sub-groups to which they belong. That first-preferences model is then post-stratified to better reflect the electoral divisions used at the 2019 Australian Federal Election. Finally, these post-stratified first-preference estimates are used as an input to a trained model that accounts for preferences distribution to estimate a two-party-preferred vote share for every electoral division.

We compare the effect of two political opinion datasets as inputs to our model. The first is a dataset from the Life in Australia survey, and the second is a dataset from the Smartvote Australia voter advice application. To be clear, we analyse these political opinion inputs entirely separately, but with the same approach and model. This allows us to compare the effect of these different datasets. In this way our paper is analogous to, albeit the inverse of, @Cohn2016 who gave the same data to a variety of pollsters to see how, given the same data, the different methods affected their estimates.

The central results of our model relate to seat counts expected by party for the 2019 Australian Federal Election. Using the Smartvote Australia dataset we expected the Australian Labor Party to lose 19 seats, win 93 seats, and for there to be 39 that were too close to call. Using the Life in Australia dataset, we expected the Australian Labor Party to lose 60 seats, win 29 seats, and for there to be 62 that were too close to call. In actual fact, the Australian Labor Party lost 83 seats and won 68 seats.

Our paper contributes in both academic and practical ways. Academically, we establish a framework for evaluating forecasting models of multi-district elections. We then develop a model for forecasting elections in Australia. Our model is electoral-division-based, but it is less influenced by outlying results than, say, electoral-division-specific polling as it shares information between electorates. Our model goes some way to accounting for preferences distribution, but it has enough flexibility to allow for differences between elections. Training our model required considerable data wrangling, including spatial and time series issues and the matching of disparate datasets. Our datasets are publicly available to make it easier for other researchers to improve on our method. Practically, we show a way in which existing political polling could be analysed in a way that is more meaningful. Our evaluation framework and the model that we put together is relevant to any multi-district system, such as Australia, Canada, and the UK, or even systems such as the US Electoral College.  Our model does not require any additional information, hence there would be no additional cost imposed to adopt our approach. Furthermore, the results obtained by our approach may help enhance the credibility of polling forecasts, as they better reflect both the realities of multi-district systems and the uncertainty inherent in any survey approach.

@@

Here we discuss the framework that we developed to characterise a useful forecasting model of parliamentary elections and how the model that we sketched out and implemented relates to each aspect of that framework.
<!-- We then examine how our results improve our understanding of the circumstances in which MRP is appropriate and  -->
<!-- Finally, we conclude mention some of the shortcomings and weaknesses of our approach and discuss areas for future development. -->

Our framework to characterise the usefulness of a forecasting model of multi-district elections has five aspects. Firstly, a model should be interpretable and provide reasonable estimates of the number of electoral divisions that each party will win. Secondly, a model should account for preferences or turnout, depending on the specifics of the election being considered. Thirdly, a model should operate in as reproducible and transparent way as is possible, and would ideally be replicable. Fourthly, a model should produce results that account for the fact that it is a probability model and not rely on the audience for those results to be statistically fluent. And finally, a model should not be worse than traditional approaches in terms of speed or cost.

@@

A reasonable model must be interpretable because it is unavoidably partially assessed on its explanatory value, as elections occur only every few years. @dowding2019prediction [p. 1,001] distinguish between 'pragmatic prediction' which is what we have engaged in here, and 'scientific prediction', and argue that '[a]ssessing how good [pragmatic] predictions are is based on how closely the assigned probabilities match what actually happens.' But given how irregular and idiosyncratic each election is, a useful forecasting model would also engage in 'scientific prediction', in which the explanatory value of the model is also important. Although substantial improvements to a useful forecasting model of multi-district elections would occur after each election, an interpretable model allows continuous smaller-scale improvements. <!--Additionally, one reason that political opinion polling occurs is so that campaigns and the media can use it. An uninterpretable model would be limiting in this regard. -->

Probabilistic outputs mean that uncertainty is an inseparable aspect of the results. This is important when the occurrence of anything other than the central estimate would have significant effects; hence critical in a multi-district elections setting. The shift to probabilistic results has occurred in a variety of fields and scenarios, for instance, the United Nations moved toward probabilistic population projections in recent years [@raftery2014bayesian].

@@

While no model in this area will ever be entirely accurate ex post, a reasonable model would only be inaccurate for discernable reasons. For instance, perhaps the assumptions around preferences turned out to be inappropriate. A model that was wrong for reasons that cannot be determined is difficult to update in an appropriate way.

If a model does not explicitly output electoral division counts by party then it is being done implicitly, and this would firstly, be unlikely to appropriately propagate uncertainty, and secondly, not be reproducible or transparent, which are concerns touched on in the third aspect of our framework.

@@

The second aspect of our framework for evaluating the reasonableness of a model for forecasting multi-district elections is that is accounts for preferences, or turnout, as appropriate, in a way that propagates uncertainty.

Australia's election system follows a preferenctial voting system, in which an instant runoff election is held for each election. Estimating the distribution of preference flows from eliminated candidates is crucial to predicting the results Australian elections. It is easy to do this in a deterministic way, for instance, by assuming that what happened last election happens again, or to assume some other reasonable sounding but similar approach. We argue that a useful forecasting model would account for preferences in a probabilistic way.

The most important reason for this is the propagation of uncertainty. If preferences are not probabilistically modelled in some way, then the resulting forecasts are likely to be too confident in the central estimates. One way to see whether this is likely to be done is whether the estimates of the two-party-preferred shares are more distributed than the first-preferences shares.

@@

The issues around preferences in Australia are relevant to turnout in many other multi-district elections. Turnout is less of an issue in Australia, due to compulsory voting, and many of the studies of actual changes, as opposed to survey responses, are necessarily focused on elections from over a century ago, such as @fowler2013electoral. Using survey data, @dassonneville2019compulsory find that around 60 per cent of survey respondents would have voted even in the absence of compulsory voting. @mackerras1999compulsory [p. 217] find compulsory voting reduces the Coalition vote by around five per cent, although @jackman1999non argues it is not this large. In any case, our model is trained on datasets for which compulsory voting has consistently applied.

@@

@jamieson2019signaling [p. 19,233] describe transparency '[a]s a prerequisite to scrutiny' and argue that it allows 'others to examine study design, execution, and data and, as a result, replicate results.' However, we go further than @jamieson2019signaling and argue that a useful forecasting model would not just require '[a]rchiving data and analysis plans in publicly available repositories', but would also impose a timing and access constraint. Specifically, the data and analysis of a useful forecasting model of multi-district elections must be available at the same time as any public release relating to the forecasting model, be that an academic research paper or press coverage, and it must be just as easily accessed. <!--For instance, it is not enough that datasets and analysis be available on request, if the coverage of the forecasting model is in the media; they must be able to be accessed as easily as the media, for instance in a publicly available repository.-->

@@

Further, as poll-of-poll models increased in popularity, analysts increasingly began to report overall estimates in terms of probabilities. But firstly, in an Australian context, the overall result of an election is essentially a Bernoulli random variable, and so claiming success when a party wins where the  model gave that party a probability of 0.2 is disingenuous. Even if this wasn't the case, there is considerable evidence that it is extremely difficult to correctly interpret probabilities [@handmer2007communicating], and that this can have an effect on electoral behaviour [@westwood2019projecting].

While the tendancy to downplay uncertainty is understandable given the constraints faced by journalists [@ricketson2019like] and the difficulty that even experts have appreciating the nuances of statistical concepts [@hoekstra2018improving; @greenland2016statistical], the results of a useful forecasting model should not require statistical fluency to understand. Misunderstandings on the part of consumers of estimates are, to a large extent, the responsibility of those who produce estimates. To make the impact of uncertainty clear, communication of a model should explicitly account for areas of uncertainty in layman's terms by, for example, reporting districts that are too close to call.

@@

@@

Our main data requirement is individual-level political opinion polling data that elicits an individual's first-preference, along with demographic and geographic information about that individual such as their electoral division. We use datasets from the Life in Australia survey and the Smartvote Australia voter advice application. In order to post-stratify our estimates we require nested electoral-division-level data for those same demographic and geographic variables. We obtain those data from the 2016 Australian Census. Finally, in order to train the preferences model we require data from past elections on preference flows, which we obtain from the Australian Electoral Commission. We detail these datasets in this section. Our data wrangling was completed in the statistical language R [@RCitation2018], using the broom [@broomreference], haven [@havenreference], foreign [@foreignreference], rgdal [@rgdalreference], sp [@spreferenceone; @spreferencetwo], and tidyverse [@tidyversereference] packages.

@@

The Australian Federal Parliament is divided into two chambers and in this paper we focus on the 'House of Representatives', which is the lower house and from where government is formed. In 2019 for the purposes of lower house representation Australia was divided into 151 electoral divisions each of which with one representative. The boundaries of the electoral divisions are determined by an independent commission that is meant to ensure that each electoral division contains roughly the same amount of people, given various historical constraints such as not crossing state boundaries, and reasonableness. For the 2019 election the number of electors in each electoral division ranged from 69,332 in Solomon which is in the Northern Territory, to 124,507 in Cowper which is in New South Wales. The median was 109,430 electors, while the mean was 108,770 electors, and the standard deviation was 9,686. The level of gerrymandering in Australia is generally considered to be low [@ComaLago2018]. Although there have been infamous episodes historically, these tended to be at a state level, rather than the federal level which is the focus of this paper.

@@

There are typically five to ten candidates in each electoral division, and voters express preferences over candidates (see Figure \ref{fig:ballotexample} for an example of a completed ballot paper). A voter's most-preferred candidate is their 'first-preference', and they rank every candidate through to their least-preferred candidate. The initial step to decide the electoral division's representative is to count first-preferences. The candidate with the least number of first-preference votes is eliminated, and the second-preferences of the voters whose first-preference was eliminated are distributed to the remaining candidates. This preferences distribution continues until there are only two candidates. The vote proportions at this point are 'two-candidate-preferred' proportions, and the winner is the candidate with a two-candidate-preferred percentage that is at least 50 per cent.[^marginsfootnote]

@@

[^moreyearsfootnote]: Data on preferences are available since preferential voting was introduced in 1918, for instance via the AustralianElections R package [@AustralianElections]. Nonetheless we train the model on four elections' worth of data. This is a balance between over-fitting by using too few elections, and ensuring the relevance of the preferences that we are training on. Additionally, because of the need to convert the results into the electoral divisions used in the 2019 Federal Election  every additional election that we train the preferences model on would introduce more error.

The main difficulty with training this model is that the boundaries of the electoral divisions can change between elections. This means that, for instance, as shown in Figure \ref{fig:map}, constructed using @ggmapreference, the Division of Canberra as it was in the 2016 election is not geographically identical to the Division of Canberra as it was for the 2019 election. In order to account for this we obtain both first preferences and two-party-preferred historical counts on a voting-booth basis. Using the latitude and longitude of each booth, and maps of the electoral divisions as they were at the 2019 election, we assign each booth to the appropriate 2019 electoral division. This allows us to construct a geographically consistent dataset that we use to train this level of our model.
<!-- [^noprepollfootnote] -->

@@

In Australia, the Australian Electoral Commission sends teams to prisons, hospitals, and other remote locations in an effort to provide an opportunity to vote. As part of our process we will not be counting these votes. The exact number that we remove, each year, is summarised in Table \ref{tab:hospitals}.

@@

hospitals %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        digits = 2, 
        format.args = list(big.mark = ","),
        caption = "Number of first preference votes dropped when the electoral divisions are made consistent",
        linesep = ""
        )  
# %>% 
#   kable_styling(font_size = 8)

@@

Over the elections from 2007 to 2016 the median share of first-preferences for the Greens was between six and ten per cent to nine per cent. However, in 2004 their maximum share of first-preferences was 20 per cent, and that maximum increased each election to be competitive in 2010 when they first won a seat in the lower house, and had reached 43 per cent by 2016.

There has consistently been a few independent and other parties that receive a competitive share of first-preferences. This has included Tony Windsor, and Bob Katter in 2007; Robert Oakeshott in 2010; Andrew Wilkie, Cathy McGowen and Clive Palmer in 2013; and most recently, Rebekha Sharkie in 2016. But in general, independent and other parties do not get a large share of first-preference votes.

The lowest average percentage of first-preference votes that the Labor party received over this time was 31 per cent in 2013, and the highest was 42 per cent in 2007. Over this time the Liberal party received a higher average in four of the five elections with a low of 39 per cent in 2016 and a high of 43 in 2004. The only election where the Labor party had a higher share of the first-preference votes was 2007 when Kevin Rudd replaced John Howard as prime minister. The largest difference was in 2013 when the Liberal party had an average share that was larger than the average share for the Labor party by almost 11 percentage points. The National party has a competitive median share of first-preference votes at between 29 and 44 per cent. However, they have a large spread, possibly given the nature of the electoral divisions in which they run

@@

```{r firstpreftotwoprefs, fig.cap = "First-preferences compared with two-party-preferred, by election", include = TRUE, echo = FALSE}
knitr::include_graphics(here::here("outputs/figures/twoPP_vs_primary_consistentdivision.pdf"))
```

Preferences tend to flow toward the major parties for the simple reason that the smaller parties and independents are usually eliminated first. Labor tends to be above a linear best-fit line and the Liberals tend to be below it. To be elected requires a post-preferences proportion of more than 0.5, and so the implication is that for a given first-preference proportion, all else being equal, it is more likely that a Labor candidate will be elected than a Liberal candidate.

@@

Table \ref{tab:bothummarystats} shows the distribution of support by major party grouping. LNP refers to any of the Liberal, National, or Liberal National (in the case of Queensland) parties. ALP is the Australian Labor Party, GRN is the Greens, Other accounts for any other party.

>>>>

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "Forecasting Multi-District Elections"
thanks: "We thank the Social Research Centre, and in particular Ben Phillips, for graciously providing data. Thank you to Monica Alexander, Lauren Kennedy, John McAndrews, Benjamin Stevens, Michael Donnelly, as well as seminar participants at the University of Toronto Political Behaviour Group and the 2019 POP AusPSA Standing Group Annual Workshop for their assistance and helpful suggestions. Our code and datasets are available at: https://github.com/RohanAlexander/ForecastingMultiDistrictElections. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca. For the latest version please click [here](https://github.com/RohanAlexander/ForecastingMultiDistrictElections/raw/master/outputs/paper/paper.pdf)."
author:
- name: Rohan Alexander
  affiliation: University of Toronto
- name: Patrick Dumont
  affiliation: Australian National University
- name: Patrick Leslie
  affiliation: Australian National University
abstract: "The overall outcome of an election in a multi-district system turns on the number of electoral divisions won by each party. Despite this, traditional political opinion polls tend to focus on broad national-level estimates of support for major parties. In this paper we develop a framework to characterise a useful forecasting model of multi-district elections. We then implement an approach to forecast the number of electoral divisions won by each of the larger political parties in Australia at the 2019 Federal Election. Our approach first uses survey data and multi-level regression with post-stratification (MRP) to estimate electoral-division-level first-preference shares. To account for Australia's use of preferential voting, we then apply a model that we trained on previous election results to estimate electoral-division-level two-party-preferred shares. We find that our approach performs well out-of-sample, and it has additional advantages over traditional polling including improved interpretability, transparency, and better communication of statistical uncertainty. Our framework allows the consistent evaluation of forecasting models of elections in multi-district systems. Finally, our paper also improves our understanding of the circumstances in which MRP is appropriate. Specifically, we find that despite the bias of voter advice applications, and despite the relatively small number of respondents per electoral division for a typical political opinion poll, both can still provide informative results on an electoral-division basis."
keywords: "Multi-district elections, political opinion polls, forecasting, multi-level regression with post-stratification, Australia."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: references.bib
biblio-style: "apalike"
endnote: no

@@

In this paper we show that it is possible to use the individual-level responses already collected by political opinion polling firms, along with publicly available datasets and accessible quantitative methods to produce estimates of the number of electoral divisions won by each party. We first establish a framework for evaluating a forecasting model for multi-district elections. It consists of five aspects. These are that a model should: be interpretable and provide reasonable estimates of the number of electoral divisions that each party will win; account for preferences and/or turnout; operate in a reproducible and transparent way; produce results that account for the fact that it is a probability model; and, finally, not be worse than traditional approaches in terms of speed or cost. Within this framework, the method that we implement involves first estimating the first-preference share of each party in each electoral division based on survey data using multi-level regression with post-stratification. We then apply a model that we trained to estimate division-level two-party-preferred shares.
<!-- Seat-specific anomalies can be accounted for by adding additional layers to the model, or by inputting detailed information about that seat. And the model, which is centered around multi-level regression with post-stratification goes some way to ensuring that a handful of odd results do not need to be manually adjusted or removed. -->

Multi-level regression with post-stratification (MRP) [@GelmanLittle1997] is a popular way to analyse opinion by way of providing good-quality estimates of subsets of the population. It uses a regression model to consider individual-level survey responses as being determined by various sub-groups and then rebuilds the sample to better match the population. In this way MRP can not only allow a better understanding of opinion, but also allow us to analyse data that may otherwise be unusable and consider outcomes within the context of distributions and uncertainty. MRP has been used in a variety of settings and one of its advantages is that a survey analysed in this way can be biased to a larger extent than normal because the approach should go some way to account for this. For instance, in the US @WangRothschildGoelGelman2015 used data collected from the XBox platform, which tends to be both younger and more male than the US electorate, and were able to draw meaningful insights about the 2012 US Presidential Election.

In this paper we construct a Bayesian MRP model to generate division-level two-party-preferred estimates for the Australian 2019 Federal Election. In our regression model an individual's first-preference is a function of the demographic and geographic sub-groups to which they belong. That first-preferences model is then post-stratified based on census data to better reflect the electoral divisions used at the 2019 Australian Federal Election. Finally, these post-stratified first-preference estimates are used as an input to a model trained on four past elections' worth of preferences distributions to estimate a two-party-preferred vote share for every electoral division.

We compare the effect of two political opinion datasets as inputs to our model. The first is a dataset from the Life in Australia survey, and the second is a dataset from the Smartvote Australia voter advice application. To be clear, we analyse these political opinion inputs entirely separately, but with the same approach and model. This allows us to compare the effect of these different datasets. In this way our paper is analogous to, albeit the inverse of, @Cohn2016 who gave the same data to a variety of pollsters to see how, given the same data, their different methods affected their estimates.

The central results of our model relate to expected seat counts by party for the 2019 Australian Federal Election. Using the Smartvote Australia dataset we expected the Australian Labor Party to lose 19 seats, win 93 seats, and for there to be 39 that were too close to call. Using the Life in Australia dataset, we expected the Australian Labor Party to lose 60 seats, win 29 seats, and for there to be 62 that were too close to call. In actual fact, the Australian Labor Party lost 83 seats and won 68 seats.

Our paper contributes in both academic and practical ways. Academically, we establish a framework for evaluating forecasting models of multi-district elections. We then develop a model for forecasting elections in Australia. Our model is electoral-division-based, but it is less influenced by outlying results than, say, electoral-division-specific polling as it shares information between electorates. Our model goes some way to accounting for preferences distribution, but it has enough flexibility to allow for differences between elections. Training our model required considerable data wrangling, including dealing with spatial and time series issues and the matching of disparate datasets. Our datasets are publicly available to make it easier for other researchers to improve on our method. Practically, we show a way in which existing political polling could be analysed in a way that is more meaningful than the usual approach. Our evaluation framework and the model that we put together is relevant to any multi-district system, such as Australia, Canada, and the UK, or even systems such as the US Electoral College.  Our model does not require any additional survey information, hence there would be no additional cost imposed to adopt our approach. Furthermore, the results obtained by our approach may help enhance the credibility of political opinion polling, as they better reflect both the realities of multi-district systems and the uncertainty inherent in any survey approach.

@@

Here we develop a framework to characterise a useful forecasting model of multi-district elections. It has five aspects. Firstly, a model should be interpretable and provide reasonable estimates of the number of electoral divisions that each party will win. Secondly, a model should account for preferences or turnout, depending on the specifics of the election being considered. Thirdly, a model should operate in as reproducible and transparent way as is possible, and would ideally be replicable. Fourthly, a model should produce results that account for the fact that it is a probability model and not rely on the audience for those results to be statistically fluent. And finally, a model should not be worse than traditional approaches in terms of speed or cost.

@@

A reasonable model must be interpretable because it is unavoidably partially assessed on its explanatory value, as elections typically occur only every few years. @dowding2019prediction [p. 1,001] distinguish between 'pragmatic prediction' which is what we have engaged in here, and 'scientific prediction', and argue that '[a]ssessing how good [pragmatic] predictions are is based on how closely the assigned probabilities match what actually happens.' But given how irregular and idiosyncratic each election is, a useful forecasting model would also engage in 'scientific prediction', in which the explanatory value of the model is also important. Although substantial improvements to a useful forecasting model of multi-district elections would occur after each election, an interpretable model allows continuous smaller-scale improvements over an election cycle. <!--Additionally, one reason that political opinion polling occurs is so that campaigns and the media can use it. An uninterpretable model would be limiting in this regard. -->

Probabilistic outputs mean that uncertainty ought to be an inseparable aspect of the results. This is important when the occurrence of anything other than the central estimate would have significant effects; hence critical in a multi-district election setting. The shift to probabilistic results has occurred in a variety of fields and scenarios, for instance, the United Nations moved toward probabilistic population projections in recent years [@raftery2014bayesian]. There is no reason that it should not place a larger role in political opinion polling.

@@

While no model in this area will ever be entirely accurate ex post, a reasonable model would only be inaccurate for discernable reasons. For instance, perhaps the assumptions around preferences turned out to be wrong. A model that was wrong for reasons that could not be determined would be difficult to update in an appropriate way.

If a model does not explicitly output electoral division counts by party then will be done implicitly. But this would firstly, be unlikely to appropriately propagate uncertainty, and secondly, not be reproducible or transparent, which are concerns touched on in the third aspect of our framework.

@@

The second aspect of our framework for evaluating the reasonableness of a model for forecasting multi-district elections is that it accounts for preferences, or turnout, as appropriate, in a way that propagates uncertainty.

Australia's federal electoral system follows a preferential voting system, in which an instant runoff election is held for each election. This was first introduced in a by-election in 1918 largely in an attempt to obtain political gains [@reilly2001preferential, p. 85]. Estimating the distribution of preference flows from eliminated candidates is crucial to predicting the results of Australian elections.

It is straight-forward to account for preferences in a deterministic way, for instance, by assuming that what happened last election happens again, or to assume some other reasonable sounding but similar approach. We argue that a useful forecasting model would account for preferences in a probabilistic way. The most important reason for this is the propagation of uncertainty. If preferences are not probabilistically modelled in some way, then the resulting forecasts are likely to be too confident in their central estimates. One way to tell whether this has likely been done is whether the estimates of the two-party-preferred shares are more widely distributed than the first-preferences shares.

@@

The issues around preferences in Australia are relevant to turnout in many other multi-district elections. Turnout is less of an issue in Australia, due to compulsory voting, and many of the studies of actual changes, as opposed to survey responses, in its absence are necessarily focused on elections from over a century ago, such as @fowler2013electoral. Using survey data, @dassonneville2019compulsory find that around 60 per cent of survey respondents would have voted even in the absence of compulsory voting. @mackerras1999compulsory [p. 217] find compulsory voting reduces the Coalition vote by around five per cent, although @jackman1999non argues it is not this large. In any case, our model is trained on datasets for which compulsory voting has consistently applied. @McNair1983 describe further some of the features of Australia in terms of political opinion polling and compare Australia with other countries.

@@

@jamieson2019signaling [p. 19,233] describe transparency '[a]s a prerequisite to scrutiny' and argue that it allows 'others to examine study design, execution, and data and, as a result, replicate results.' However, we go further than @jamieson2019signaling and argue that a useful forecasting model would not just require '[a]rchiving data and analysis plans in publicly available repositories', but would also impose a timing and access constraint. Specifically, the data and analysis of a useful forecasting model of multi-district elections must be available at the same time as any public release relating to the forecasting model, be that an academic research paper or press coverage, and it ought to be just as easily accessed. <!--For instance, it is not enough that datasets and analysis be available on request, if the coverage of the forecasting model is in the media; they must be able to be accessed as easily as the media, for instance in a publicly available repository.-->

@@

Further, recently, as poll-of-poll models have increased in popularity, analysts have increasingly reported overall estimates in terms of probabilities. But firstly, in an Australian context, the overall result of an election is essentially a Bernoulli random variable, and so claiming success when a party wins where the model gave that party a probability of 0.2 is disingenuous. And even if this was not the case, there is considerable evidence that it is difficult to correctly interpret probabilities [@handmer2007communicating], and that this could have an effect on electoral behaviour [@westwood2019projecting].

While the tendency to downplay uncertainty is understandable given the constraints faced by journalists [@ricketson2019like] and the difficulty that even experts have appreciating the nuances of statistical concepts [@hoekstra2018improving; @greenland2016statistical], the results of a useful forecasting model should not require statistical fluency to understand. Misunderstandings on the part of consumers of estimates are, to a large extent, the responsibility of those who produce estimates. To make the impact of uncertainty clear, communication of model results should explicitly account for areas of uncertainty in layman's terms by, for example, reporting which electoral districts are too close to call.

@@

Ultimately, ours is not just an academic exercise. Political opinion polls have come in for much scrutiny and criticism recently and there are many who are dissatisfied with the status quo. For instance, the major Nine-owned newspapers, The Age and The Sydney Morning Herald, paused their use of polling after the 2019 election. While there is some commercial incentive for political opinion pollsters to move to better methods as political opinion polling is a way for them to attract business, there are also considerable commercial realities in the way of adopting new approaches. Hence, the minimal requirement of an alternative is that it not impose greater financial cost or require additional time, than what is currently being done.

@@

Our main data requirement is individual-level political opinion polling data that elicits an individual's first-preference, along with demographic and geographic information about that individual such as their electoral division. We use datasets from the Life in Australia survey and the Smartvote Australia voter advice application. In order to post-stratify our estimates we require nested electoral-division-level data for those same demographic variables. We obtain these data from the 2016 Australian Census. Finally, in order to train the preferences model we require data from past elections on preference flows, which we obtain from the Australian Electoral Commission. We detail these datasets in this section. Our data wrangling was completed in the statistical language R [@RCitation2018], using the broom [@broomreference], haven [@havenreference], foreign [@foreignreference], rgdal [@rgdalreference], sp [@spreferenceone; @spreferencetwo], and tidyverse [@tidyversereference] packages.

@@

The Australian Federal Parliament is divided into two chambers and in this paper we focus on the 'House of Representatives', which is the lower house and from where government is formed. In 2019 for the purposes of lower house representation Australia was divided into 151 electoral divisions, each with one representative. The boundaries of the electoral divisions are determined by an independent commission that is meant to ensure that each electoral division contains roughly the same amount of electors, given various historical constraints such as not crossing state boundaries, and reasonableness. For the 2019 election the number of electors in each electoral division ranged from 69,332 in Solomon which is in the Northern Territory, to 124,507 in Cowper which is in New South Wales. The median was 109,430 electors, while the mean was 108,770 electors, and the standard deviation was 9,686. The level of gerrymandering in Australia is generally considered to be low [@ComaLago2018]. Although there have been infamous episodes historically, these tended to be at a state level, rather than the federal level which is the focus of this paper.

@@

There are typically five to ten candidates in each electoral division, and voters express preferences over candidates (see Figure \ref{fig:ballotexample} for an example of a completed ballot paper). A voter's most-preferred candidate is their 'first-preference', and they rank every candidate through to their least-preferred candidate. The initial step to decide the electoral division's representative is to count first-preferences. The candidate with the least number of first-preference votes is eliminated, and the second-preferences of the voters whose first-preference was eliminated are distributed to the remaining candidates. This preferences distribution continues until there are only two candidates. The vote proportions at this point are 'two-candidate-preferred' proportions, and that electoral division's representative will be the candidate with a two-candidate-preferred percentage that is at least 50 per cent.[^marginsfootnote]

@@

[^moreyearsfootnote]: Data on preferences are available since preferential voting was introduced in 1918, for instance via the AustralianElections R package [@AustralianElections] or the eechidna R package [@eechidnacite]. Nonetheless we train the model on four elections' worth of data. This is a balance between over-fitting by using too few elections, and ensuring the relevance of the preferences that we are training on. Additionally, because of the need to convert the results into the electoral divisions used in the 2019 Federal Election  every additional election that we train the preferences model on introduces more error.

The main difficulty with training this model is that the boundaries of the electoral divisions can change between elections. This means that, for instance, as shown in Figure \ref{fig:map}, constructed using the ggmap package of @ggmapreference, the Division of Canberra as it was in the 2016 election is not geographically identical to the Division of Canberra as it was for the 2019 election. In order to account for this we obtain both first preferences and two-party-preferred historical counts on a voting-booth basis. Using the latitude and longitude of each booth, and maps of the electoral divisions as they were at the 2019 election, we assign each booth to the appropriate 2019 electoral division. This allows us to construct a geographically consistent dataset that we use to train this level of our model.
<!-- [^noprepollfootnote] -->

@@

In Australia, the Australian Electoral Commission sends teams to prisons, hospitals, and other remote locations in an effort to provide an opportunity to vote. Our process means that we will not be counting these votes. The exact number that we remove, each election, is summarised in Table \ref{tab:hospitals}.

@@

hospitals %>%
  kable(format = "latex", 
        booktabs = TRUE, 
        digits = 2, 
        format.args = list(big.mark = ","),
        caption = "Number of first preference votes dropped when the electoral divisions are made consistent.",
        linesep = ""
        )  
# %>% 
#   kable_styling(font_size = 8)

@@

Over the four elections from 2007 to 2016 the median share of first-preferences for the Greens was between six and ten per cent to nine per cent. However, in 2004 their maximum share of first-preferences was 20 per cent, and that maximum increased each election to be competitive in 2010 when they first won a seat in the lower house, and had reached 43 per cent by 2016.

There have consistently been a few independent and other parties that receive a competitive share of first-preferences. This has included Tony Windsor and Bob Katter in 2007; Robert Oakeshott in 2010; Andrew Wilkie, Cathy McGowen, and Clive Palmer in 2013; and, Rebekha Sharkie in 2016. But in general, independent and other parties do not get a large share of first-preference votes.

The lowest average percentage of first-preference votes that the Labor party received over this time was 31 per cent in 2013, and the highest was 42 per cent in 2007. Over this time the Liberal Party received a higher average in all but one election with a low of 39 per cent in 2016. The only election where the Australian Labor Party had a higher share of the first-preference votes was 2007 when Kevin Rudd replaced John Howard as prime minister. The largest difference was in 2013 when the Liberal Party had an average share that was larger than the average share for the Labor party by almost 11 percentage points. The National Party has a competitive median share of first-preference votes at between 29 and 44 per cent. However, they have a large distribution, possibly given the nature of the electoral divisions in which they run

@@

```{r firstpreftotwoprefs, fig.cap = "First-preferences compared with two-party-preferred, by election.", include = TRUE, echo = FALSE}
knitr::include_graphics(here::here("outputs/figures/twoPP_vs_primary_consistentdivision.pdf"))
```

Preferences tend to flow toward the major parties for the simple reason that the smaller parties and independents are usually eliminated earlier. The Australian Labor Party tends to be above a linear best-fit line and the Liberal Party tends to be below it. To be elected requires a post-preferences proportion of more than 0.5, and so the implication is that for a given first-preference proportion, all else being equal, it is more likely that an Australian Labor Party candidate will be elected than a Liberal Party candidate.

@@

Table \ref{tab:bothummarystats} shows the distribution of support by major party grouping. LNP refers to any of the Liberal, National, or Liberal National (in the case of Queensland) parties. ALP is the Australian Labor Party, GRN is the Greens, and Other accounts for any other party.
